FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-06-12 paper.tex
Add notes on tracking down Inf, add CSTG diagrams

\subsection{Catching NaN kills}

@@

\begin{verbatim}
…
-                 FloatTracker/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
…
\end{verbatim}

We can see that the NaN appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.

Thus, with little effort we were able to track down where the NaNs were originating.
More work is needed to follow the $\infty$ back to wherever \emph{it} came from.
Fortunately, we will be able to extend FloatTracker to handle more than just NaNs---watching and logging \texttt{Inf}s should be completely tractable.
See~\cref{discussion} for more on future plans around this.

% TODO: if we get some handling with Inf working, then we should keep tracing this back.
% Where did the $\infty$ come from?

>>>>

\subsection{Catching a NaN Killer}

@@

\begin{verbatim}
…
-                 FT/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
…
\end{verbatim}

We can see that the NaN appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
Thus, with little effort we were able to track down where the NaNs were originating.
Now the question is: where did the $\infty$ come from?

FloatTracker is not limited to NaNs---we can track any kind of value that we want.
We have built-in support for tracking NaN and $\infty$; we can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)

\begin{verbatim}
check_error([-1.515f31, 2])  FT/TrackedFloat.jl:10
^(::TrackedFloat32, ::Int64) FT/TrackedFloat.jl:138
…
materialize(^)               broadcast.jl:860
top-level getproperty(P, :u) examples/sw_nan_tf.jl:14
…
\end{verbatim}

It looks like the $-\infty$ came from $-1.515e31^2$---no wonder a Float32 type couldn't handle something that large.
We can tell that because the first line shows us the arguments\footnote{There is a configuration option that allows us to control this behavior.} and the second line shows us what function got called. Moreover, we get the top-level line: we can see that it came from line 14 in our ShallowWaters file.

Tracking down why that value got so large is beyond the scope of FloatTracker right now.
Most likely we will need some domain-expertise to find a solution to this problem.
The point is, we were able to quickly and automatically arrive at this excellent starting point with FloatTracker.

\subsubsection{CSTGs help us get a bigger picture}

Running FloatTracker can produce a lot of logs.
To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
% FIXME: cite CSTG

We can produce a CSTG of both the Inf as well as the NaN logs.
Here is what the two graphs look like:\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}

\begin{figure}
  \centering
  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
  \caption{CSTG of NaN generation logs from ShallowWaters.}
  \label{fig:sw_nan_cstg}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
  \caption{CSTG of Inf generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
  \label{fig:sw_inf_cstg}
\end{figure}
